---
title: "ESM 204 HW 3"
output: html_document
date: 'Anna Zauner'
---

```{r setup, include=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, options(scipen = 999))
library(tidyverse)
library(here)
library(janitor)
library(thematic)
library(scales)
library(equatiomatic)
thematic::thematic_rmd()
thematic::thematic_on()
```


## Introduction

```{r}
# Read in the data 
data1 <- read_csv(here("data","HW3_data.csv")) %>% 
  clean_names() %>% 
  select(-x1)
```


The Biden Administration’s “interim” value is $51 per metric ton of CO2. In this analysis, we consider the distributional consequences of a household electricity tax based on the Social Cost of Carbon (SCC) to address the climate change problem.

Assumptions:
- Consumers can separated into two income groups: “high” and “low.” 
- Price ($) and quantity (kWh) estimates of demand per month for the two groups. 
- Initially, there is no tax on electricity consumption.
- The current electricity price (without any taxes) is $.10 per kWh.
- The marginal cost of producing a kWh of electricity is linear and has a price-intercept of 0.


### **Question #1**

Calculating MEC

Assumptions:
- One kWh of electricity emits 0.85 pounds of CO2 
- Interim value of $51 per metric ton of CO2 

```{r}
# PIVOT LONGER WIH INCOME AS VARIABLE 

data_long <- data1 %>% 
  pivot_longer(cols = c(q_low_kwh, q_high_kwh),
               names_to = 'income',
               values_to = 'kwh') %>% 
  mutate(income = case_when(income == 'q_low_kwh' ~ 'low',
                   income == 'q_high_kwh' ~ 'high'))

# RUN CALCULATION FOR MEC

price_ton <- 51

# METRIC TON UNIT CONVERSION

price_lb <- price_ton/2205

price_lb

# Multiply by 0.85 to get MEC/kWh

price_kWh <- price_lb*0.85

price_kWh

```

Based on the calculations, the MEC per kWh of electricity is **0.0197**.

git test


### **Question #2**

Calculating:

a. Aggregate monthly demand for electricity
b. Supply curve for electricity 
c. Benefit to consumers and producers under status quo
d. Environmental cost under status quo


```{r}



```

### **Question #3**

Calculating consumer benefit divide between "high" and "low" income consumers.

```{r}

```


### **Question #4**

Deriving the optimal electricity tax (cents per kWh) with interim SCC. 
Assumptions:
- Low income group experience disproportionate share of impacts from climate change
- Climate externality is borne entirely by "low" income group

Calculating effects of tax on:

a. Amount of electricity produced and consumed
b. Price of electricity
c. Welfare of "high" income consumers
d. Welfare of "low" income consumers

```{r}

```


### **Question #5**

Assumptions:
- All tax revenue will be distributed to consumers in proportion to their pre-tax consumption
- True SCC may be much higher than $51 (range: $51, $75, $100, $125, $150 per metric ton of CO2)

Calculating effects of SCC-based electricity tax for:

a. Welfare of "high" income consumers
b. Welfare of "low" income consumers
c. Electricity producers




```{r}

```



### **Question #6**


```{r}

```

